---
title: "Differntial Evolution with DEoptim"
output: html_notebook
---

illustrate the package’s usage with the minimization of the Rastrigin function in R2 which is a common test for global optimization: 
uncomment to call stored intermediate populations that allows us examine the progress of optimization in detail
```{r}
#Install DEoptim package 
#library(DEoptim)
#Rastrigin <- function(x)(sum(x^2 - 10 * cos(2 * pi * x)) + 20)
set.seed(1234)
#DEoptim(fn = Rastrigin,lower = c(-5, -5), upper = c(5, 5), control = list(storepopfrom = 1))
```

Risk Allocation portfolios
determining what risk measue to use
compute the log-return series and the mean and covariance matrix estimators
#to define tickers
```{r}
library("quantmod")
tickers <- c("GE", "IBM", "JPM", "MSFT", "WMT")
getSymbols(tickers,
from = "2000-12-01",
to = "2010-12-31")
P <- NULL
for(ticker in tickers) {
tmp <- Cl(to.monthly(eval(parse(text = ticker))))
P <- cbind(P, tmp)
}
colnames(P) <- tickers
R <- diff(log(P))
R <- R[-1,]
mu <- colMeans(R)
sigma <- cov(R)

```
compute equal weight portfolio
to see how much each assest contributes to the total portfolio
```{r}
library("PerformanceAnalytics")
pContribCVaR <- ES(weights = rep(0.2, 5),
                     method = "gaussian",
                     portfolio_method = "component",
                     mu = mu,
                     sigma = sigma)$pct_contrib_ES
rbind(tickers, round(100 * pContribCVaR, 2))
round(100 * mu , 2)
round(100 * diag(sigma)^(1/2), 2)

```

Let's solve using the DEoptim package
first define objective function to minimize
```{r}
set.seed(1234)
obj <- function(w) {
if (sum(w) == 0) 
{w <- w + 1e-2}
w <- w / sum(w)
CVaR <- ES(weights = w,
method = 'gaussian',
portfolio_method = 'component',
mu = mu, sigma = sigma)
tmp1 <- CVaR$ES
tmp2 <- max(CVaR$pct_contrib_ES - 0.225, 0)
out <- tmp1 + 1e3 * tmp2
}
set.seed(1234)
out <- DEoptim(fn = obj,
               lower = rep(0, 5), 
               upper = rep(1, 5))
out$optim$bestval
wstar <- out$optim$bestmem
wstar <- wstar / sum(wstar)
rbind(tickers, round(100 * wstar, 2))
100 * (sum(wstar * mu) - mean(mu))
```


lets compare to optim
```{r}
out <- optim(par = rep(0.2, 5),
               fn = obj,
               method = "L-BFGS-B",
               lower = rep(0, 5),
               upper = rep(1, 5))
out$value
```
compare to nlminb
```{r}
out <- nlminb(start = rep(0.2, 5),
                objective = obj,
                lower = rep(0, 5),
                upper = rep(1, 5))
out$objective

```



minimizing the largest CVaR contribution subject to a return target can beimplemented as:
```{r}
obj <- function(w) {
  if(sum(w) == 0) { w <- w + 1e-2 }
  w <- w / sum(w)
  contribCVaR <- ES(weights = w,
             method = "gaussian",
             portfolio_method = "component",
             mu = mu,
            sigma = sigma)$contribution
  tmp1 <- max(contribCVaR)
  tmp2 <- max(mean(mu) - sum(w * mu), 0)
  out <- tmp1 + 1e3 * tmp2
}
set.seed(1234)
out <- DEoptim(fn = obj,
                 lower = rep(0, 5),
                  upper = rep(1, 5))
wstar <- out$optim$bestmem
wstar <- wstar / sum(wstar)
rbind(tickers, round(100 * wstar, 2))
100 * (sum(wstar * mu) - mean(mu))
```

